---
title: "DataCleaning"
author: "Amanda Pennino"
date: "4/2/2021"
output: html_document
---


New columns to be added:
#`Measurement` (Analyte, Conductivity, Salinity, Hardness, etc.)
#`MeasurementNotes`: Any descriptor that was in orginal `Focal Salts/Ions` column.
#`TempSingleValue/Min`, `TempMax`, `TempPlusMinus`: Altered temperatures

```{r echo = FALSE}
library(tidyverse)

```

```{r echo = FALSE}

#Most recent download of Raw Salt Data Sheet (as of 05/20/21)
DATAraw <- read_csv("data/raw_salt_data_052021.csv") 

#Give each line of data a reference id
DATAraw$DataID <- 1:nrow(DATAraw)

#Select columns of interest, keep `DataID` for table joining later.

DATA <- DATAraw %>%
  select(DataID, RefID, Reviewer, Salt_Iontype_Conductivity, Units_salt_concentration_original, Minimum_concentration_original, Maximum_concentration_original, `Rate of change in salt (if applicaple)`, Temperature)

```


New column for measurements, cleaning the `Focal: Salt/Ion type (if known) OR Conductivity` column for spelling and capitalization errors.
```{r}

DATA$Measurement = DATA$Salt_Iontype_Conductivity


#### Calcium ion, calcium chloride, calcium carbonates
DATA$Measurement <- ifelse(DATA$Measurement %in% c('calcium chloride', 'CaCl', "CaCL2", 'calcium chloride (CaCl2.2H2O)'), "CaCl2", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('Calcium', 'Ca', 'Ca++', 'calcium'), "Ca2+", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('calcium carbonate', 'Calcium carbonate'), "CaCO3", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('CaCL2-NaHCO3' ), "CaCl2-NaHCO3", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('CaCL2-Na2SO4' ), "CaCl2-Na2SO4", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('CaCL2-NaCl' ), "CaCl2-NaCl", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('CaCL2-KCl' ), "CaCl2-KCl", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('CaCL2-K2SO4' ), "CaCl2-K2SO4", DATA$Measurement)


#### Chloride ion and other chlorides
DATA$Measurement <- ifelse(DATA$Measurement %in% c('chloride', 'Cl', 'Chloride' ), "Cl-", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('sodium chloride', 'nacl', "NaCL", 'Sodium chloride'), "NaCl", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('magnesium chloride', 'MgCl', 'MgCL', 'MgCL2', 'Magnesium Chloride', 'Magnesium chloride'), "MgCl2", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCl2-CaCL2','MgCL2-CaCl2') , "MgCl2-CaCl2", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCL2-CaSO4') , "MgCl2-CaSO4", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCL2-K2SO4') , "MgCl2-K2SO4", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCL2-NaCl') , "MgCl2-NaCl", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCL2-KCl') , "MgCl2-KCl", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCL2-NaHCO3') , "MgCl2-NaHCO3", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCL2-KHCO3') , "MgCl2-KHCO3", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCL2-Na2SO4') , "MgCl2-Na2SO4", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('CaCl2-NA2SO4 ') , "CaCl2-Na2SO4 ", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('KCL'), "KCl", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('ammonium chloride'), "NH4Cl", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('ammonium sulfate'), "(NH4)2SO4", DATA$Measurement)

#### Sodium ion
DATA$Measurement <- ifelse(DATA$Measurement %in% c('sodium', 'Sodium', 'Na'), "Na+", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('sodium bicarbonate'), "NaHCO3", DATA$Measurement)


#### Potassium
DATA$Measurement <- ifelse(DATA$Measurement %in% c('potassium', 'K', 'Potassium'), "K+", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('potassium bicarbonate'), "KHCO3", DATA$Measurement)


### Sulfates
DATA$Measurement <- ifelse(DATA$Measurement %in% c('sulfate', 'SO4', 'SO42'), "SO4 2-", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('NA2SO4', 'Na2SO4 (sodium sulfate)','sodium sulfate' ), "Na2SO4", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgCL2-NaHCO3') , "MgCl2-NaHCO3", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgSO4-CaCL2') , "MgSO4-CaCl2", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgSO4-NA2SO4') , "MgSO4-Na2SO4", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('K2SO4-NA2SO4') , "K2SO4-Na2SO4", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('CaSO4-NA2SO4') , "CaSO4-Na2SO4", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('MgSO4-MgCL2') , "MgSO4-MgCl2", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('CaSO4-CaCL2') , "CaSO4-CaCl2", DATA$Measurement)

#### Nitrogen
DATA$Measurement <- ifelse(DATA$Measurement %in% c('nitrate', "Nitrate" ), "NO3-", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('ammonium', "Ammonium" ), "NH4+", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('ammonia'), "NH3+", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('sodium nitrate'), "NaNO3", DATA$Measurement)


#### Magnesium
DATA$Measurement <- ifelse(DATA$Measurement %in% c('magnesium', "Magnesium", "Mg" ), "Mg3+", DATA$Measurement)

#### Conductance, salinity, Hardness, Dissolved Solids
DATA$Measurement <- ifelse(DATA$Measurement %in% c('conductivity', 'electrical conductivity', 'Electrical Conductivity', 'Conducitivity', 'conductivity (EC)', 'Electrical Conductance', 'Electric Conductance'), "Conductivity", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('salinity', 'salinty' ), "Salinity", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('hardness'), "Hardness", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('total dissolved solids', 'TDS'), "Total Dissolved Solids", DATA$Measurement)


DATA$Measurement <- ifelse(DATA$Measurement %in% c('Specific conductance'), "Specific Conductance", DATA$Measurement)


#### Water types
DATA$Measurement <- ifelse(DATA$Measurement %in% c('artificial seawater', 'synthetic sea water', 'Artificial seawater'), "Artifical Seawater", DATA$Measurement)
DATA$Measurement <- ifelse(DATA$Measurement %in% c('seawater', 'Seawater', 'natural sea water', 'natural seawater'), "Natural Seawater", DATA$Measurement)

DATA$Measurement <- ifelse(DATA$Measurement %in% c('Bicarbonate'), "HCO3-", DATA$Measurement)

#unique(DATA$Measurement)

```



Take out additional information in `Measurement` and add it to a `MeasurementNotes` column. 
```{r}
DATA$MeasurementNotes <- NA

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (SO4:HCO3 ratio)")] <- "SO4:HCO3 ratio"
DATA$Measurement[DATA$Measurement == "Specific Conductance (SO4:HCO3 ratio)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: KCl)")] <- "Focal Salt: KCl"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: KCl)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: MgCl2)")] <- "Focal Salt: MgCl2"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: MgCl2)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance( Focal Salt: NaHCO3)")] <- "Focal Salt: NaHCO3"
DATA$Measurement[DATA$Measurement == "Specific Conductance( Focal Salt: NaHCO3)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: seasalt)")] <- "seasalt"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: seasalt)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: CaCO3)")] <- "Focal Salt: CaCO3"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: CaCO3)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: Na2SO4)")] <- "Focal Salt: Na2SO4"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: Na2SO4)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: NaNO3)")] <- "Focal Salt: NaNO3"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: NaNO3)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: Mg3+)")] <- "Focal Salt: Mg3+"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: Mg3+)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: NH3+)")] <- "Focal Salt: NH3+"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: NH3+)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: PO4)")] <- "Focal Salt: PO4"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: PO4)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: Cl-)")] <- "Focal Salt: Cl-"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: Cl-)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: iron oxide)")] <- "Focal Salt: iron oxide"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: iron oxide)"] <- "Specific Conductance"
                                          
DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: Ca Na SO4 dominant (simulated river water))")] <- "Focal Salt: Ca Na SO4 dominant (simulated river water)"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: Ca Na SO4 dominant (simulated river water))"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: Ca Mg HCO3 SO4 dominant)")] <- "Focal Salt: Ca Mg HCO3 SO4 dominante"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: Ca Mg HCO3 SO4 dominant)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salts: Mg Ca SO4 (Mg = 2x Ca mg/L) (Winding Shoals water))")] <- "Focal Salts: Mg Ca SO4 (Mg = 2x Ca mg/L) (Winding Shoals water))"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salts: Mg Ca SO4 (Mg = 2x Ca mg/L) (Winding Shoals water))"] <- "Specific Conductance"     

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Dominant ions are Ca Mg SO4 HCO3)")] <- "Dominant ions are Ca Mg SO4 HCO3"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Dominant ions are Ca Mg SO4 HCO3)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salts: Ca Mg SO4 HCO3)")] <- "Focal Salts: Ca Mg SO4 HCO3"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salts: Ca Mg SO4 HCO3)"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Specific conductance (Dominant ions are Ca Mg SO4 HCO3)")] <- "Dominant ions are Ca Mg SO4 HCO3"
DATA$Measurement[DATA$Measurement == "Specific conductance (Dominant ions are Ca Mg SO4 HCO3)"] <- "Specific Conductance"
                             
DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salts: Ca Mg SO4 HCO3 dominant)")] <- "Focal Salts: Ca Mg SO4 HCO3 dominant"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salts: Ca Mg SO4 HCO3 dominant)"] <- "Specific Conductance"                               
DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: Ca2+)")] <- "Focal Salt: Ca2+"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: Ca2+)"] <- "Specific Conductance"                                  
DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: Na, K, HCO3 are dominant)")] <- "Focal Salt: Na, K, HCO3 are dominant+"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: Na, K, HCO3 are dominant)"] <- "Specific Conductance"          
DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: NaCl & KCl)")] <- "Focal Salt: NaCl & KCl"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: NaCl & KCl)"] <- "Specific Conductance" 

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: SO4, HCO5)")] <- "Focal Salt: SO4, HCO5"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: SO4, HCO5)"] <- "Specific Conductance" 

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: SO42)")] <- "Focal Salt: SO42"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: SO42)"] <- "Specific Conductance" 

DATA$MeasurementNotes[which(DATA$Measurement == "Specific Conductance (Focal Salt: NaCl)")] <- "Focal Salt: NaCl"
DATA$Measurement[DATA$Measurement == "Specific Conductance (Focal Salt: NaCl)"] <- "Specific Conductance" 

DATA$Measurement[DATA$Measurement == "Specific Conductance \n"] <- "Specific Conductance"

DATA$MeasurementNotes[which(DATA$Measurement == "Ca (applied as CaCO3)")] <- "Applied as CaCO3"
DATA$Measurement[DATA$Measurement == "Ca (applied as CaCO3)"] <- "Ca2+"

DATA$MeasurementNotes[which(DATA$Measurement == "Ca (applied as sulfates; and in combination with Mg)")] <- "Applied as sulfates; and in combination with Mg"
DATA$Measurement[DATA$Measurement == "Ca (applied as sulfates; and in combination with Mg)"] <- "Ca2+"

DATA$MeasurementNotes[which(DATA$Measurement == "Ca (as component of seawater solution)")] <- "Ca as component of seawater solution)"
DATA$Measurement[DATA$Measurement == "Ca (as component of seawater solution)"] <- "Ca2+"

DATA$MeasurementNotes[which(DATA$Measurement == "chloride; but study also focused on nitrate, potassium, magnesium, sulfate")] <- "chloride; but study also focused on nitrate, potassium, magnesium, sulfate"
DATA$Measurement[DATA$Measurement == "chloride; but study also focused on nitrate, potassium, magnesium, sulfate"] <- "Cl-"

DATA$MeasurementNotes[which(DATA$Measurement == "Cl (as NaCl)")] <- "As NaCl"
DATA$Measurement[DATA$Measurement == "Cl (as NaCl)"] <- "Cl-"

DATA$MeasurementNotes[which(DATA$Measurement == "conductivity (NaCl added)")] <- "NaCl added"
DATA$Measurement[DATA$Measurement == "conductivity (NaCl added)"] <- "Conductivity"

DATA$MeasurementNotes[which(DATA$Measurement == "conductivity (potassium)")] <- "Potassium"
DATA$Measurement[DATA$Measurement == "conductivity (potassium)"] <- "Conductivity"

DATA$MeasurementNotes[which(DATA$Measurement == "conductivity (SO4:HCO3 ratio)")] <- "conductivity (SO4:HCO3 ratio)"
DATA$Measurement[DATA$Measurement == "conductivity (SO4:HCO3 ratio)"] <- "Conductivity"

DATA$MeasurementNotes[which(DATA$Measurement == "conductivity (sodium)")] <- "conductivity (sodium)"
DATA$Measurement[DATA$Measurement == "conductivity (sodium)"] <- "Conductivity"

DATA$MeasurementNotes[which(DATA$Measurement == "conductivity: Ca Mg HCO3 SO4 dominant")] <- "conductivity: Ca Mg HCO3 SO4 dominant"
DATA$Measurement[DATA$Measurement == "conductivity: Ca Mg HCO3 SO4 dominant"] <- "Conductivity"

DATA$MeasurementNotes[which(DATA$Measurement == "KCl (in water with 109 mg/L as CaCO3 hardness)")] <- "In water with 109 mg/L as CaCO3 hardness"
DATA$Measurement[DATA$Measurement == "KCl (in water with 109 mg/L as CaCO3 hardness)"] <- "KCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 103 mg/L as CaCO3 hardness) (\"100 hard\")")] <- "In water with 103 mg/L as CaCO3 hardness) (\"100 hard\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 103 mg/L as CaCO3 hardness) (\"100 hard\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 109 mg/L as CaCO3 hardness) (\"100 hard\")")] <- "In water with 109 mg/L as CaCO3 hardness) (\"100 hard\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 109 mg/L as CaCO3 hardness) (\"100 hard\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 193 mg/L as CaCO3 hardness) (\"200 hard\")")] <- "In water with 193 mg/L as CaCO3 hardness) (\"200 hard\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 193 mg/L as CaCO3 hardness) (\"200 hard\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 204 mg/L as CaCO3 hardness) (\"200 hard\")")] <- "NaCl (in water with 204 mg/L as CaCO3 hardness) (\"200 hard\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 204 mg/L as CaCO3 hardness) (\"200 hard\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 265 mg/L as CaCO3 hardness) (\"300 hard\")")] <- "NaCl (in water with 265 mg/L as CaCO3 hardness) (\"300 hard\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 265 mg/L as CaCO3 hardness) (\"300 hard\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 299 mg/L as CaCO3 hardness) (\"300 hard\")")] <- "NaCl (in water with 299 mg/L as CaCO3 hardness) (\"300 hard\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 299 mg/L as CaCO3 hardness) (\"300 hard\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 43 mg/L as CaCO3 hardness) (\"50 hard\")")] <- "NaCl (in water with 43 mg/L as CaCO3 hardness) (\"50 hard\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 43 mg/L as CaCO3 hardness) (\"50 hard\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 61 mg/L as CaCO3 hardness) (\"50 hard\")")] <- "NaCl (in water with 61 mg/L as CaCO3 hardness) (\"50 hard\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 61 mg/L as CaCO3 hardness) (\"50 hard\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (in water with 87 mg/L as CaCO3 hardness) (\"MHRW\")")] <- "NaCl (in water with 87 mg/L as CaCO3 hardness) (\"MHRW\")"
DATA$Measurement[DATA$Measurement == "NaCl (in water with 87 mg/L as CaCO3 hardness) (\"MHRW\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl (with treatment concentrations measured as conductivity)")] <- "NaCl (with treatment concentrations measured as conductivity)"
DATA$Measurement[DATA$Measurement == "NaCl (with treatment concentrations measured as conductivity)"] <- "NaCl"



DATA$MeasurementNotes[which(DATA$Measurement == "NaCl spiked with Ca (in water with 130 mg/L as CaCO3 hardness) (\"50 hard + Ca\")")] <- "NaCl spiked with Ca (in water with 130 mg/L as CaCO3 hardness) (\"50 hard + Ca\")"
DATA$Measurement[DATA$Measurement == "NaCl spiked with Ca (in water with 130 mg/L as CaCO3 hardness) (\"50 hard + Ca\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl spiked with HCO3 (in water with 40 mg/L as CaCO3 hardness) (\"50 hard _ HCO3\")")] <- "NaCl spiked with HCO3 (in water with 40 mg/L as CaCO3 hardness) (\"50 hard _ HCO3\")"
DATA$Measurement[DATA$Measurement == "NaCl spiked with HCO3 (in water with 40 mg/L as CaCO3 hardness) (\"50 hard _ HCO3\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl spiked with K (in water with 41 mg/L as CaCO3 hardness) (\"50 hard + K\")")] <- "NaCl spiked with K (in water with 41 mg/L as CaCO3 hardness) (\"50 hard + K\")"
DATA$Measurement[DATA$Measurement == "NaCl spiked with K (in water with 41 mg/L as CaCO3 hardness) (\"50 hard + K\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "NaCl spiked with SO4 (in water with 41 mg/L as CaCO3 hardness) (\"50 hard + SO4\")")] <- "NaCl spiked with SO4 (in water with 41 mg/L as CaCO3 hardness) (\"50 hard + SO4\")"
DATA$Measurement[DATA$Measurement == "NaCl spiked with SO4 (in water with 41 mg/L as CaCO3 hardness) (\"50 hard + SO4\")"] <- "NaCl"


DATA$MeasurementNotes[which(DATA$Measurement == "Salinity (commercial salt)")] <- "Salinity (commercial salt)"
DATA$Measurement[DATA$Measurement == "Salinity (commercial salt)"] <- "Salinity"


DATA$MeasurementNotes[which(DATA$Measurement == "salinity (dominated by Ca and Na)")] <- "Dominated by Ca and Na"
DATA$Measurement[DATA$Measurement == "salinity (dominated by Ca and Na)"] <- "Salinity"


DATA$MeasurementNotes[which(DATA$Measurement == "salinity (Road Runner Ice Melt)")] <- "Road Runner Ice Melt"
DATA$Measurement[DATA$Measurement == "salinity (Road Runner Ice Melt)"] <- "Salinity"



DATA$MeasurementNotes[which(DATA$Measurement == "Specific conductance (Dominant ions are Ca Mg SO4 HCO3)")] <- "Dominant ions are Ca Mg SO4 HCO3"
DATA$Measurement[DATA$Measurement == "Specific conductance (Dominant ions are Ca Mg SO4 HCO3)"] <- "Specific Conductance"


DATA$MeasurementNotes[which(DATA$Measurement == "Na (as component of seawater solution)")] <- "As a component of seawater solution"
DATA$Measurement[DATA$Measurement == "Na (as component of seawater solution)"] <- "Na+"



DATA$MeasurementNotes[which(DATA$Measurement == "conductivity (synthetic sea salt,\nOcean Nature)")] <- "Synthetic sea salt. Ocean Nature"
DATA$Measurement[DATA$Measurement == "conductivity (synthetic sea salt,\nOcean Nature)"] <- "Conductivity"


DATA$MeasurementNotes[which(DATA$Measurement == "conductivity: Ca Na SO4 dominant (simulated river water)")] <- "Ca Na SO4 dominant (simulated river water)"
DATA$Measurement[DATA$Measurement == "conductivity: Ca Na SO4 dominant (simulated river water)"] <- "Conductivity"


DATA$MeasurementNotes[which(DATA$Measurement == "Mg (as component of seawater solution)")] <- "As a component of seawater solution"
DATA$Measurement[DATA$Measurement == "Mg (as component of seawater solution)"] <- "Mg3+"

DATA$MeasurementNotes[which(DATA$Measurement == "K (as component of seawater solution)")] <- "As a component of seawater solution"
DATA$Measurement[DATA$Measurement == "K (as component of seawater solution)"] <- "K+"


DATA$MeasurementNotes[which(DATA$Measurement == "salinity (natural seawater)")] <- "Natural Seawater"
DATA$Measurement[DATA$Measurement == "salinity (natural seawater)"] <- "Salinity"

#unique(DATA$Measurement)

```

Cleaning temperature data and creating new columns. 
```{r}

DATA$TempNotes = NA

DATA$Temperature <- ifelse(DATA$Temperature %in% c("na", "not given", "Not given", "not mentioned","not provided", "not reported","Not specified", "not stated", "Not stated", "? (ASTM standard for mussel tox)"), NA, DATA$Temperature)

DATA$TempNotes[which(DATA$Temperature == "Not stated, but mussels were stored at 22 +/1 2 deg C")] <- "Mussels were stored at 22 +/1 2 deg C"
DATA$Temperature[DATA$Temperature == "Not stated, but mussels were stored at 22 +/1 2 deg C"] <- NA

DATA$TempNotes[which(DATA$Temperature == "7.5, 10, 12.5, 15, 20, 20.5 C")] <- "7.5, 10, 12.5, 15, 20, 20.5 C"
DATA$TempNotes[which(DATA$Temperature == "5, 7.5, 10, 12.5, 15, 20, 20.5 C")] <- "5, 7.5, 10, 12.5, 15, 20, 20.5 C"
DATA$TempNotes[which(DATA$Temperature == "5, 10, 12.5, 15, 20, 20.5 C")] <- "5, 10, 12.5, 15, 20, 20.5 C"
DATA$TempNotes[which(DATA$Temperature == "20, 24, 28, 32C")] <- "20, 24, 28, 32C"
DATA$TempNotes[which(DATA$Temperature == "15.1, 14.9, 11.1")] <- "15.1, 14.9, 11.1" 
DATA$TempNotes[which(DATA$Temperature == "10, 12.5, 15, 20, 20.5 C" )] <- "10, 12.5, 15, 20, 20.5 C"  

DATA$TempNotes[which(DATA$Temperature == "varied among streams" )] <- "varied among streams" 

DATA$Temperature <- ifelse(DATA$Temperature %in% c("7.5, 10, 12.5, 15, 20, 20.5 C", "5, 7.5, 10, 12.5, 15, 20, 20.5 C", "5, 10, 12.5, 15, 20, 20.5 C", "20, 24, 28, 32C", "10, 12.5, 15, 20, 20.5 C", "15.1, 14.9, 11.1", "varied among streams"), "Multiple, See TempNotes", DATA$Temperature)

DATA$Temperature <- ifelse(DATA$Temperature %in% c("2 degrees- 10 degrees"), "2-10", DATA$Temperature)

DATA$TempNotes[which(DATA$Temperature == "16 - 32 (ambient range)" )] <- "ambient range" 
DATA$Temperature <- ifelse(DATA$Temperature %in% c("16 - 32 (ambient range)" ), "16-32", DATA$Temperature)

DATA$TempNotes[which(DATA$Temperature == "2 - 30 (temp was also an experimental variable)" )] <- "temp was also an experimental variable"
DATA$Temperature <- ifelse(DATA$Temperature %in% c("2 - 30 (temp was also an experimental variable)"), "2-30", DATA$Temperature)



#### Create SingleValue/Min & Max columns
DATA <- DATA %>%
  separate(Temperature, c("TempSingleValue/Min", "TempMax"), sep = "-", remove = FALSE)


#### Create a Plus/Minus temperature column
DATA <- DATA %>%
   mutate(TempPlusMinus = sapply(strsplit(`TempSingleValue/Min`, "±"), "[", 2))


#### Clean up TempSingleValue/Min
DATA <- DATA %>%
  mutate(`TempSingleValue/Min` = sapply(strsplit(`TempSingleValue/Min`, "±"), "[", 1))

DATA$`TempSingleValue/Min` <- str_remove(DATA$`TempSingleValue/Min`, "[(+]")
DATA$`TempSingleValue/Min` <- str_remove(DATA$`TempSingleValue/Min`, "[+/]")
DATA$`TempSingleValue/Min` <- ifelse(DATA$`TempSingleValue/Min` %in% c("16 C"), "16", DATA$`TempSingleValue/Min`)
DATA$`TempSingleValue/Min` <- ifelse(DATA$`TempSingleValue/Min` %in% c("14°C"), "14", DATA$`TempSingleValue/Min`)
DATA$`TempSingleValue/Min` <- ifelse(DATA$`TempSingleValue/Min` %in% c("23 C"), "23", DATA$`TempSingleValue/Min`)
DATA$`TempSingleValue/Min` <- ifelse(DATA$`TempSingleValue/Min` %in% c("ambiant", "ambient"), "Ambient", DATA$`TempSingleValue/Min`)
DATA$`TempSingleValue/Min` <- ifelse(DATA$`TempSingleValue/Min` %in% c("25 ", "25 conductivity corrected for 25)"), "25", DATA$`TempSingleValue/Min`)



#### Clean up TempMax
DATA$TempMax <- str_remove(DATA$TempMax, "[C]")


```


Cleaning units
```{r}

DATA$SaltUnits = DATA$Units_salt_concentration_original


DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("Percent NaCl", "percent", "percent SW", "%sw", "% seawater"), "%", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("g L-1", "g/l", "g NaCl/L", "g L-", "gm/litre"), "g/L", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("mg Ca /L", "mg Ca/L", "mg CaCO3/L", "mg Cl / L", "mg Cl- / L", "mg Cl/L", "mg K / L", "mg l", "mg Mg/L", "mg SO4/L", "mg/ L", "mg/l", "mg/L CaCO3"), "mg/L", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("us/cm", "uS/cm", "uScm-1", "μS/cm", "µS/cm", "μScm–1" ), "µS/cm", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("mmol Ca/L", "mmol Mg/L", "mmol l-1", "mmol K/L", "mmol Na/L", "mMol/L", "mmole/L", "mM/L", "m mol l-1"), "mmol/L", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("mS cm ^-1", "mS cm−1", "mS / cm", "ms/cm"), "mS/cm", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("parts per thousand (ppt)", "per mil", "ppt NaCl (LC50)"), "ppt", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("ugl-1", "ug/L"), "µg/L", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("umho/cm", "umohs/cm", "umhos/cm", "umohs", "umhos", "μMho", "micro-mhos"), "µmho/cm", DATA$SaltUnits)

#DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("mhmos/m"), "mmhos/m", DATA$SaltUnits)


DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("not given", "not provided"), "Not provided", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("mOsM"), "mOsm", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("mOsm*kg/l"), "mOsm/kg", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("millimole"), "mmol", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("mM/l", "mM/L", "mM"), "mmol/L", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("um/cm"), "µm/cm", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("uS/m"), "µS/m", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("umol/L", "µM", "uM Ca2+", "uM"), "µmol/L", DATA$SaltUnits)

DATA$SaltUnits <- ifelse(DATA$SaltUnits %in% c("psu (practical salinity units)"), "psu", DATA$SaltUnits)

DATA$SaltUnits[DATA$SaltUnits == "see \"Notes on effect\""] <- "not provided"


#unique(sort(DATA$SaltUnits))

```

Output the most recent version of Corrections
```{r}
#reorder columns
DATA<- DATA[,c(1,2,3,4,12,5,16,6,7,8,9,10,11,15,13,14)]

write_csv(DATA, "data/RawSalts_Clean.csv")

```
