---
title: "Plotting"
author: "Stephen and Amanda"
date: "5/27/2021"
output: html_document
---

```{r echo = FALSE}

library(tidyverse)

```

```{r}
DATA <- read_csv("data/fulldatabase.csv")



unique(DATA$Taxonomic_classification)
unique(DATA$study_location)
unique(DATA$Measurement)

DATA$Taxonomic_classification <- ifelse(DATA$Taxonomic_classification %in% c("mollusca", "Mollusca", "moluska"), "Mollusca", DATA$Taxonomic_classification)

DATA$study_location <- ifelse(DATA$study_location %in% c("Field", "field", "Fied"), "Field", DATA$study_location)
DATA$study_location <- ifelse(DATA$study_location %in% c("Lab", "lab"), "Lab", DATA$study_location)

cond_ranges<-DATA %>%
  filter(Taxonomic_classification %in% c("amphibian","Crustacean", "Fish", "Insect", "Mollusca", "non-insect, non-arthropod", "Zooplankton")) %>%
  filter(!is.na(Max_Conductivity_uScm)) %>%
  filter(!is.na(Min_Conductivity_uScm)) %>%
  group_by(Taxonomic_classification, study_location) %>%
  summarise(Min_Conductivity_uScm = min(Min_Conductivity_uScm), Max_Conductivity_uScm = max(Max_Conductivity_uScm))
  
ion_ranges<-DATA %>%
  filter(Taxonomic_classification %in% c("amphibian","Crustacean", "Fish", "Insect", "Mollusca", "non-insect, non-arthropod", "Zooplankton")) %>%
  filter(!is.na(Max_Concentration_ppm)) %>%
  filter(!is.na(Min_Concentration_ppm)) %>%
  group_by(Taxonomic_classification, study_location) %>%
  summarise(Min_Concentration_ppm = min(Min_Concentration_ppm), Max_Concentration_ppm = max(Max_Concentration_ppm))
  
  write_csv(cond_ranges, file = "data/condranges.csv")
  write_csv(ion_ranges, file = "data/ionranges.csv")


```




```{r}


GeomSplitViolin <- ggproto("GeomSplitViolin", GeomViolin, draw_group = function(self, data, ..., draw_quantiles = NULL){
  data <- transform(data, xminv = x - violinwidth * (x - xmin), xmaxv = x + violinwidth * (xmax - x))
  grp <- data[1,'group']
  newdata <- plyr::arrange(transform(data, x = if(grp%%2==1) xminv else xmaxv), if(grp%%2==1) y else -y)
  newdata <- rbind(newdata[1, ], newdata, newdata[nrow(newdata), ], newdata[1, ])
  newdata[c(1,nrow(newdata)-1,nrow(newdata)), 'x'] <- round(newdata[1, 'x']) 
  if (length(draw_quantiles) > 0 & !scales::zero_range(range(data$y))) {
    stopifnot(all(draw_quantiles >= 0), all(draw_quantiles <= 
                                              1))
    quantiles <- ggplot2:::create_quantile_segment_frame(data, draw_quantiles)
    aesthetics <- data[rep(1, nrow(quantiles)), setdiff(names(data), c("x", "y")), drop = FALSE]
    aesthetics$alpha <- rep(1, nrow(quantiles))
    both <- cbind(quantiles, aesthetics)
    quantile_grob <- GeomPath$draw_panel(both, ...)
    ggplot2:::ggname("geom_split_violin", grid::grobTree(GeomPolygon$draw_panel(newdata, ...), quantile_grob))
  }
  else {
    ggplot2:::ggname("geom_split_violin", GeomPolygon$draw_panel(newdata, ...))
  }
})

geom_split_violin <- function (mapping = NULL, data = NULL, stat = "ydensity", position = "identity", ..., draw_quantiles = NULL, trim = TRUE, scale = "area", na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) {
  layer(data = data, mapping = mapping, stat = stat, geom = GeomSplitViolin, position = position, show.legend = show.legend, inherit.aes = inherit.aes, params = list(trim = trim, scale = scale, draw_quantiles = draw_quantiles, na.rm = na.rm, ...))
}

set.seed(20160229)

my_data = data.frame(
  y=c(rnorm(1000), rnorm(1000, 0.5), rnorm(1000, 1), rnorm(1000, 
1.5)),
  x=c(rep('a', 2000), rep('b', 2000)),
  m=c(rep('i', 1000), rep('j', 2000), rep('i', 1000))
)

#Needs to be long data

#Measurement #Concentration #Min/Max

TEST <- DATA %>% select(Measurement, Min_Concentration_ppm, Max_Concentration_ppm)


TEST <- TEST %>% pivot_longer(-Measurement, names_to = "MinMax", values_to = "Concentration")


TEST %>% 
filter(Measurement %in% c("Ca2+", "Cl-")) %>%
ggplot(aes(Measurement, Concentration, fill=MinMax)) +
  geom_split_violin(trim = TRUE) +
  geom_boxplot(width = 0.10, notch = FALSE, notchwidth = .4, outlier.shape = NA, coef=0) +
  labs(x="Analye",y="Concentration") +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  scale_x_discrete(labels=c("0" = "Control\nCondition", "1" = "GM\nCondition")) +
  scale_fill_manual(values=c("blue", "green"))



# DATA %>% 
# filter(Measurement %in% c("Cl-")) %>%
#   ggplot(aes(Min_Concentration_ppm, Max_Concentration_ppm, fill= Measurement)) + 
#   geom_split_violin(trim = TRUE) + 
#   geom_boxplot(width = 0.25, notch = FALSE, notchwidth = .4, outlier.shape = NA, coef=0) +
#   labs(y = "Concentration", x = "Analyte") +
#   theme_classic() +
#   theme(text = element_text(size = 12)) 
#   #scale_x_discrete(labels=c("0" = "Control\nCondition", "1" = "GM\nCondition")) +
#   #scale_fill_manual(values=c("#E69F00", "#999999"), 
#                     #name="Survey\nPart",
#                     #breaks=c("1", "2"),
#                     #labels=c("CaCO3", "Ca2+"))


```

```


Going to work on ion stuff here.

Seawater salinity = 35,000 ppm
```{r}


DATA %>% 
  filter(Measurement %in% c("CaCO3","Ca2+", "NaHCO3", "NaCl", "KCl", "Cl-", "K+", "Na+", "MgCl2")) %>%
  ggplot(aes(x = Measurement, y = Min_Concentration_ppm))+
  geom_boxplot()+
  theme_classic()


DATA %>% 
  filter(Measurement %in% c("CaCO3","Ca2+", "NaHCO3", "NaCl", "KCl", "Cl-", "K+", "Na+", "MgCl2")) %>%
  ggplot(aes(x = Measurement, y = Max_Concentration_ppm))+
  geom_violin(aes(colour = "#1268FF"), alpha = 0.5)+
  geom_violin(aes(y = Min_Concentration_ppm, colour = "#3268FF"), alpha = 0.5)+
  geom_jitter(height = 0.1, width = 0.2, alpha = 0.1)+
  scale_y_continuous(trans='log10')+
  theme_classic()

#length(DATA$Measurement[DATA$Measurement == "HCO3-"])                                     
                                           
unique(DATA$Measurement)


```



```{r}

#Lets remove extreme values for now

DATA %>% 
  filter(Measurement %in% c("Specific Conductance","Salinity", "Total Dissolved Solids")) %>%
  mutate(Min_Conductivity_uScm = ifelse(Min_Conductivity_uScm > 10000, NA, Min_Conductivity_uScm),
              Max_Conductivity_uScm = ifelse(Max_Conductivity_uScm > 60000, NA, Max_Conductivity_uScm))%>%
  ggplot(aes(x = Measurement, y = Max_Conductivity_uScm))+
  geom_violin(aes(fill = "#1268FF"), alpha = 0.4)+
  geom_violin(aes(y = Min_Conductivity_uScm, fill = "#3268FF"), alpha = 0.4)+
  geom_jitter(height = 0.1, width = 0.3, alpha = 0.1)+
  scale_y_continuous(trans='log10')+
  theme_classic()

#Without extreme values removed
DATA %>% 
  filter(Taxonomic_classification %in% c("amphibian","Crustacean", "Fish", "Insect", "mollusca", "non-insect, non-arthropod", "Zooplankton")) %>%
  ggplot(aes(x = Taxonomic_classification, y = Max_Conductivity_uScm))+
  geom_violin(aes(fill = "#1268FF"), alpha = 0.3)+
  geom_violin(aes(y = Min_Conductivity_uScm, fill = "#3268FF"), alpha = 0.3)+
  geom_jitter(height = 0, width = 0.3, alpha = 0.1)+
  scale_y_continuous(trans='log10')+
  theme_classic()
  


```


```{r}

unique(DATA$Taxonomic_classification)
unique(DATA$study_location)
unique(DATA$Measurement)

DATA$Taxonomic_classification <- ifelse(DATA$Taxonomic_classification %in% c("mollusca", "Mollusca", "moluska"), "Mollusca", DATA$Taxonomic_classification)

DATA$study_location <- ifelse(DATA$study_location %in% c("Field", "field", "Fied"), "Field", DATA$study_location)
DATA$study_location <- ifelse(DATA$study_location %in% c("Lab", "lab"), "Lab", DATA$study_location)

cond_ranges<-DATA %>%
  filter(Taxonomic_classification %in% c("amphibian","Crustacean", "Fish", "Insect", "Mollusca", "non-insect, non-arthropod", "Zooplankton")) %>%
  filter(!is.na(Max_Conductivity_uScm)) %>%
  filter(!is.na(Min_Conductivity_uScm)) %>%
  group_by(Taxonomic_classification, study_location) %>%
  summarise(min = min(Min_Conductivity_uScm), max = max(Max_Conductivity_uScm))
  
ion_ranges<-DATA %>%
  filter(Taxonomic_classification %in% c("amphibian","Crustacean", "Fish", "Insect", "Mollusca", "non-insect, non-arthropod", "Zooplankton")) %>%
  filter(!is.na(Max_Concentration_ppm)) %>%
  filter(!is.na(Min_Concentration_ppm)) %>%
  group_by(study_location, Taxonomic_classification) %>%
  summarise(min = min(Min_Concentration_ppm), max = max(Max_Concentration_ppm))
  
  write_csv(cond_ranges, file = "cond_ranges.csv")
  write_csv(ion_ranges, file = "ion_ranges.csv")

```

